---
title: "Statistical Methods for Accounting Systems"
author: 
    - Stijn Masschelein (University of Western Australia) \newline
    - Frank Moers (Maastricht University)
date: "04/04/2019"
fontsize: 14pt
header-includes: 
  - \definecolor{eqfocus}{HTML}{EB811B}
output: 
  binb::metropolis:
    fig_caption: no
  # beamer_presentation:
  #   theme: "metropolis"
  #   keep_tex: true
# classoption: "aspectratio=169"
bibliography: tex/complement.bib
---

<!--
We are revising the paper and mostly focusing on making the contribution
clearer.
-->

# Introduction

## Complementarity of Accounting Practices

```{r setup, include=F}
knitr::opts_chunk$set(echo = FALSE)
```

```{r packages, include = F, warnings = F, results = 'hide'}
suppressMessages(library(tidyverse, quietly = TRUE, warn.conflicts = FALSE))
suppressMessages(library(cowplot, quietly = TRUE, warn.conflicts = FALSE))
suppressMessages(library(ggthemes, quietly = TRUE, warn.conflicts = FALSE))
suppressMessages(library(xtable, quietly = TRUE, warn.conflicts = FALSE))
theme_set(theme_cowplot(font_size = 18)
          + theme(strip.background = NULL))
```

### Management Accounting - Cross-Sectional [@Grabner2013]
  - Delegation and incentive contracts
  - Levers of control
  - Budget emphasis and budget participation

> A decision maker makes at least two **interdependent** choices.

---

### Financial Accounting
  - Audited financial reporting and voluntary disclosure [@Ball2012]

> "Because private information disclosure and audited financial
reporting are complements, their economic roles cannot be
evaluated separately."

<!--
There is a lot of literature that acknowledges the interaction
between multiple accounting choices which has implications of
doing tests.
-->

## Contingency Theory [@chenhall_management_2003; @otley_contingency_2016]

> "Researchers have attempted to explain the effectiveness of MCS
by examining designs that best suit the nature of the
environment, technology, size, structure, strategy and national
culture."

> The effectiveness of the choices depends on the **environment**

## Two Specifications

### Performance Specification

Performance = f(Incentive x Delegation, Environment)

### Demand Specification

Delegation = f(Incentive, Environment)

### Which one to use?

\footnotesize
> "Which [specification] to use is both a theoretical (assumption
of agents as perfectly or boundedly rational) and an empirical
(market imperfections, exogenous shocks, experimentation, etc)
issue because the likelihood of finding evidence of the former
depends on the latter and vice versa [@Grabner2013;
@Gerdin2004]" in @Johansson2018.

## Little guidance in the literature 

1. The performance specification works as long as a sufficient
amount of firms do not adopt the optimal strategy

2. The demand specification is vulnerable to omitted 
environmental variables

## Our conclusions

1. The performance specification has low power when a sufficient 
amount of firms avoids the worst choices.

2. Omitted environment bias when the environment affects **both**
choices.

3. Performance specification has high rate of false positives

# Theory

## Objective Function

### Unobserved Objective Function

$$
\begin{aligned}
y = & \beta_0 + (\beta_1 + \gamma_1 z + \epsilon_{1})
                  {\color{eqfocus} x_{1}}
                 + (\beta_2 + \gamma_2 z + \epsilon_{2})
                  {\color{eqfocus} x_{2}} \\
                 & + \beta_{12} {\color{eqfocus} x_{1} x_{2}}
                 - .5 \delta_1 x^2_{1}
                 - .5 \delta_2 x^2_{2} + \nu
\end{aligned}
$$

### Observable Cross-sectional Data

$$
\begin{aligned}
y &= \text{performance} \\
x_1, x_2 &= \text{management accounting practices} \\
z &= \text{environmental factor}
\end{aligned}
$$

## Specifications

\footnotesize
$$
\begin{aligned}
\mathbf{x_1} &= \beta^d_{11} + {\color{eqfocus}\beta^d_{12} \mathbf{x_2}}
    + \gamma_1^d \mathbf{z} + \mathbf{\epsilon^d} \\
\mathbf{y} &=  \beta^{p1}_0 + (\beta^{p1}_{11} + \gamma_1^{p1} \mathbf{z} )\mathbf{x_1} 
						+ (\beta_{22}^{p1} + \gamma_2^{p1} \mathbf{z} ) \mathbf{x_2} 
            + {\color{eqfocus}\beta_{12}^{p1} \mathbf{x_1} \mathbf{x_2}}
            + \delta_1^{p1} \mathbf{x^2_1} + \delta_2^{p1} \mathbf{x^2_2} 
                        + \alpha^{p1} \mathbf{z}
                        + \nu^{p1} \\
 \mathbf{y} &=  \beta^{p2}_0 + (\beta^{p2}_{11} + \gamma_1^{p2} \mathbf{z} )\mathbf{x_1} 
						+ (\beta_{22}^{p2} + \gamma_2^{p2} \mathbf{z} ) \mathbf{x_2} 
            + {\color{eqfocus}\beta_{12}^{p2} \mathbf{x_1} \mathbf{x_2}}
            + \alpha^{p2} \mathbf{z}
            + \nu^{p2} \\
 \mathbf{y} &=  \beta^{p3}_0 + \beta^{p3}_{11} \mathbf{x_1} 
						+ \beta_{22}^{p3} \mathbf{x_2} 
            + {\color{eqfocus}\beta_{12}^{p3} \mathbf{x_1} \mathbf{x_2}}
            + \alpha^{p3} \mathbf{z}
            + \nu^{p3}
\end{aligned}
$$

### Unobserved Objective Function

\scriptsize
$$
\begin{aligned}
y = & \beta_0 + (\beta_1 + \gamma_1 z + \epsilon_{1})
                  {\color{eqfocus} x_{1}}
                 + (\beta_2 + \gamma_2 z + \epsilon_{2})
                  {\color{eqfocus} x_{2}} \\
                 & + \beta_{12} {\color{eqfocus} x_{1} x_{2}}
                 - .5 \delta_1 x^2_{1}
                 - .5 \delta_2 x^2_{2} + \nu
\end{aligned}
$$

### Observable Cross-sectional Data

\scriptsize
$$
\begin{aligned}
y &= \text{performance} \\
x_1, x_2 &= \text{management accounting practices} \\
z &= \text{environmental factor}
\end{aligned}
$$

## Optimal choice

$$
\frac{\partial y}{\partial x_1} = 0 \implies
\delta_1  x_1^* = \beta_1 + \beta_{12} x_2^* 
 + \gamma_1 z + \epsilon_1
$$
$$
x_1^* = \frac{\delta_2 (\beta_{11} + \gamma_1 z + \epsilon_1) 
      + \beta_{12} (\beta_{22} + \gamma_2 z + \epsilon_2) }
     {\delta_1 \delta_2 - \beta_{12}^2}
$$

## Optimal choice

\scriptsize
$$
\frac{\partial y}{\partial x_1} = 0 \implies
{\color{eqfocus} \delta_1 } x_1^* = \beta_1 + \beta_{12} x_2^*
    + \gamma_1 z + \epsilon_1
$$
\scriptsize
$$
x_1^* = \frac{\delta_2 (\beta_{11} + \gamma_1 z + \epsilon_1) 
      + \beta_{12} (\beta_{22} + \gamma_2 z + \epsilon_2) }
     {\delta_1 \delta_2 - \beta_{12}^2}
$$

\normalsize

### Consequences

1. In equilibrium, choices determined by the environment.

2. Similar environment $\implies$ similar choices 

3. Correlations of main effects $\implies$ correlations 
of interactions $\implies$ bias when omitted environment affects 
both choices

## Statistical Issues

| Issue                   | Performance       | Demand          |
|-------------------------|-------------------|-----------------|
|Lack of Power            | Optimal           | Uninformed      |
|Multicollinearity        | Optimal           |                 |
|Omitted Environment Bias | Optimal           | Optimal         |

# Simulation

## Regressions

- Demand Specification
- Performance 1 (Theoretically Correct)
- Performance 2 (Performance 1 but no quadratic terms)
- Performance 3 (Performance 2 but no contingency interactions)

## Optimisation

- Generate environment ($z, \epsilon_i, \epsilon_2$)
- Generate O combination of choices ($x_1, x_2$)
- Pick combination with highest profit ($y$)

## Samples

- For 300 observations

  - Generate environment ($z, \epsilon_i, \epsilon_2$)
  - Generate O combination of choices ($x_1, x_2$)
  - Pick combination with highest profit ($y$)
    
- Run 4 specifications; calculate t-statistic and p-value

## Simulation

Vary parameters $\beta$ (e.g. $\beta_{12} = 0$ or $.25$), 
$\gamma$, $\delta$, $\sigma$, and $O$.

- Generate 1000 samples 

  - For 300 observations

    - Generate environment ($z, \epsilon_i, \epsilon_2$)
    - Generate O combination of choices ($x_1, x_2$)
    - Pick combination with highest profit ($y$)
      
  - Run 4 specifications; calculate t-statistic and p-value
  
- Calculate Type I error ($\beta_{12} = 0$) and power ($\beta_{12} \neq 0$)

# Results

## Main Result: No environmental factor bias

```{r main, include = FALSE}
source("R/parameters.R")
sim = readRDS("simulated_data/main_simulation.Rds")
dat = tbl_df(sim) %>%
  unnest(obs, rate) %>%
  mutate(optim = 1/rate)

dat_plot = 
  mutate(dat, 
         optim_fact = paste("O ==", optim),
         b2_fact = paste(map(b2, 1))) %>%
  filter(map(g1, 2) %in% c(0, -.33)) %>%
  arrange(optim) %>%
  mutate(optim_fact = fct_relevel(optim_fact, unique(optim_fact)))

plot = (ggplot(filter(dat_plot, map(g1, 2) == 0), 
               aes(y = stat, x = as.factor(b2_fact)))
         + facet_grid(rows = vars(label),
                      cols = vars(optim_fact),
                      labeller = label_parsed)
         + geom_hline(yintercept = tint, linetype = 3, alpha = .75)
         + geom_hline(yintercept = 0, linetype = 4, alpha = .75)
         + geom_hline(yintercept = -tint, linetype = 3, alpha = .75)
         + labs(x = expression(beta[12]), y = "t-statistic")
         + theme(strip.text.x = element_text(angle = 0, size = 16),
                 strip.text.y = element_text(angle = 0))
)
``` 

```{r plot_main0}
print(plot)
```

## Main Result: No omitted environment bias

```{r plot_main1}
print(plot + geom_tufteboxplot())
```

## Main Result: No omitted environment bias

```{r main2, include = FALSE}
plot = (ggplot(filter(dat_plot, map(g1, 2) == -0.33), 
               aes(y = stat, x = as.factor(b2_fact)))
         + facet_grid(rows = vars(label),
                      cols = vars(optim_fact),
                      labeller = label_parsed)
         + geom_hline(yintercept = tint, linetype = 3, alpha = .75)
         + geom_hline(yintercept = 0, linetype = 4, alpha = .75)
         + geom_hline(yintercept = -tint, linetype = 3, alpha = .75)
         + labs(x = expression(beta[12]), y = "t-statistic")
         + theme(strip.text.x = element_text(angle = 0, size = 16),
                 strip.text.y = element_text(angle = 0))
         + geom_tufteboxplot()
)
```

```{r plot_main2}
print(plot)
```

## Main Result: Type I error rate

```{r main_table, echo = FALSE, results='asis'}
table = dat %>%
  group_by(label,  
           b2 = unlist(map(b2, 1)), optim) %>%
    summarise(type1 = prettyNum(mean(pvalue < 0.05), dig = 2),
              power = prettyNum(mean(pvalue < 0.05 & coefficient > 0), 
                                dig = 2)) %>%
    ungroup() %>% 
    mutate(percentage = ifelse(b2 != 0, power, type1),
           statistic = ifelse(b2 != 0, "power", "type I")) %>%
    select(-c(type1, power, b2)) %>%
    spread(optim, percentage) %>%
    arrange(desc(statistic), label) %>%
    rename(specification = label)

print(xtable(filter(table),
             type = "pdf",
             label = "main-table"),
      size = "\\footnotesize",
      include.rownames = FALSE,
      sanitize.text.function = force,
      comment = FALSE
)
```

## Parameter Variations

1. Increase performance noise ($\sigma_{\nu}$)
    - The omission of quadratic terms no longer picks up the 
      demand effect.
2. Decrease marginal costs increase ($\delta$)
    - Little to no change with current parameters
3. Binary practices
    - General decrease in power

## Omitted Environment Bias

![Omitted Correlated Environmental Factor](figure-latex/spurious_plot.pdf)

## Improved Type I Errors 

### Bootstrap

```{r bootstrap, results = 'asis'}
sim = readRDS("simulated_data/boot_simulation.Rds")
dat = tbl_df(sim) %>%
  unnest(rate, obs) %>%
  mutate(optim = 1/rate)

table = dat %>%
  group_by(label, b2 = unlist(map(b2, 1)), optim) %>%
  summarise(type1 = prettyNum(mean(sign(lower_interval) 
                              == sign(higher_interval)), dig = 2),
            power = prettyNum(mean(lower_interval > 0), dig = 2)) %>%
  ungroup() %>%
  mutate(percentage = ifelse(b2 != 0, power, type1),
         statistic = ifelse(b2 != 0, "power", "type I")) %>%
  select(-c(type1, power, b2)) %>%
  spread(optim, percentage) %>%
  arrange(desc(statistic), label) %>%
  rename(specification = label)
 
print(xtable(filter(table),
             type = "pdf",
             label = "bootstrap-table"),
      size = "\\footnotesize",
      include.rownames = FALSE,
      sanitize.text.function = force,
      comment = FALSE
)
```

## Improved Type I Errors 

### Nearly Exact Covariance [@young_improved_2016]

```{r nearly-exact, results = 'asis'}
sim = tbl_df(readRDS("simulated_data/nearly_simulation.RDS",))
dat = tbl_df(sim) %>%
  unnest(rate, obs) %>%
  mutate(label = as.character(label)) %>%
  mutate(optim = 1/rate,
         label = if_else(str_detect(label, "nearly_exact"),
                         str_c(str_replace(label, "nearly_exact_", ""),
                               "corrected", sep = " "),
                         label))

table = dat %>%
  group_by(label, 
           b2 = unlist(map(b2, 1)), optim) %>%
  summarise(type1 = prettyNum(mean(pvalue < 0.05), dig = 2),
            power = prettyNum(mean(pvalue < 0.05 & coefficient > 0), 
                          dig = 2)) %>%
  ungroup() %>%
  mutate(percentage = ifelse(b2 != 0, power, type1),
         statistic = ifelse(b2 != 0, "power", "type I")) %>%
  select(-c(type1, power, b2)) %>%
  spread(optim, percentage) %>%
  arrange(desc(statistic), label) %>%
  rename(specification = label)

print(xtable(filter(table),
             type = "pdf",
             label = "nearly-table"),
      size = "\\footnotesize",
      include.rownames = FALSE,
      sanitize.text.function = force,
      comment = FALSE
)
```

# Conclusion

## Main Takeaway

- Always report demand specification

- Always control for environmental factors
  - Demand: control variables
  - Production: interaction with practices
    - Principal Component Analysis
    
- Always correct standard errors in the performance 
  specification. 

# Appendix

## General Objective Function 

### Unobserved Objective Function

$$
\begin{aligned}
y = &\beta_0 +
\Bigl(
\begin{bmatrix}
  \beta_{11} &
  \beta_{22}
\end{bmatrix}
+
z
\begin{bmatrix}
  \gamma_{1} & \gamma_{2}
\end{bmatrix}
+ \begin{bmatrix}
  \epsilon_1 & \epsilon_2
\end{bmatrix}
\Bigr)
\begin{bmatrix}
  {\color{eqfocus} x_1} \\ {\color{eqfocus} x_2}
\end{bmatrix} +
\\
& \frac{1}{2}
\begin{bmatrix}
  {\color{eqfocus} x_1} & {\color{eqfocus} x_2}
\end{bmatrix}
\begin{bmatrix}
  - \delta_1 & \beta_{12} \\
  \beta_{12} & - \delta_2
\end{bmatrix}
\begin{bmatrix}
  {\color{eqfocus} x_1} \\ {\color{eqfocus} x_2}
\end{bmatrix}
\end{aligned}
$$

### Observable Data

$$
\begin{aligned}
y &= \text{performance} \\
{\color{eqfocus} x_1, x_2} &= \text{management accounting practices} \\
z &= \text{environmental factor}
\end{aligned}
$$

## Omitted Environment

$$
\begin{aligned}
\gamma_1 \frac{cov(\mathbf{x_1 x_2}, \mathbf{x_1 z})}{var(\mathbf{x_1 z})} &= 
\gamma_1 \frac{r_{2z} + r_{12} r_{1z}}{1 + r_{1z}^2} \\
\gamma_2 \frac{cov(\mathbf{x_1 x_2}, \mathbf{x_2 z})}{var(\mathbf{x_2 z})} &= 
\gamma_2 \frac{r_{1z} + r_{12} r_{2z}}{1 + r_{2z}^2}
\end{aligned}
$$

## Matrix Notation

$$
y = \beta_0 + (\mathbf{\beta_1}^T + \mathbf{z^T} G^T + 
     \mathbf{\epsilon^T}) \mathbf{x} 
     + \frac{1}{2}\mathbf{x^T} B \mathbf{x} + \nu
$$
$$
\begin{aligned} 
    \mathbf{\beta_1} + G \mathbf{z} + \mathbf{\epsilon} & = -B \mathbf{x^*} \\
    - B^{-1} (\mathbf{\beta_1} +  G \mathbf{z} + \mathbf{\epsilon})  & = \mathbf{x^*}
\end{aligned}
$$

## Matrix: Omitted Environment Factor

$$
\begin{aligned}
    cov(\mathbf{x_1 x_2, \gamma_{im} \mathbf{z_m} \mathbf{x_i} + \gamma_{jm} \mathbf{z_m} \mathbf{x_j}} | X_{-i, -j}, Z_{-m}) = \\
    ({\color{eqfocus}\gamma^2_{im}} \zeta^p_{im} 
    + {\color{eqfocus}\gamma^2_{jm}} \zeta^p_{jm}) \lambda^p_{ij} 
    {\color{eqfocus}\beta_{ij}} + 
    {\color{eqfocus}\gamma_{im} \gamma_{jm}} (\zeta^p_{im} + \zeta^p_{jm})  
\end{aligned}
$$

## Matrix: Omitted Choice

$$
\begin{aligned}
   cov(\mathbf{x_i x_j}, \beta_{ik} \mathbf{x_k} \mathbf{x_i} + \beta_{jk} \mathbf{x_k} \mathbf{x_j}| X_{-i, -j, -k}, Z) = \\
   ({\color{eqfocus}\beta^2_{ik}} \lambda^p_{ik} 
   + {\color{eqfocus}\beta^2_{jk}} \lambda^p_{jk}) \lambda_{ij}^p
    {\color{eqfocus}\beta_{ij}} + 
    {\color{eqfocus}\beta_{ik} \beta_{jk}} (\lambda^p_{ik} +
    \lambda^p_{jk})
\end{aligned}
$$

## Second order condition 

### General condition

$$
\begin{bmatrix}
  \delta_1 &  - \beta_{12} & \dots & - \beta_{1N} \\
  - \beta_{12} &  \delta_2 & \dots & - \beta_{2N} \\
  \vdots     & \vdots     & \ddots & \vdots \\
  - \beta_{1N} & - \beta_{2N} & \dots & \delta_N
\end{bmatrix}
\text{is p.d.}
$$

### Two practice condition

$$
\delta_1 \delta_2 - \beta_{12}^2 > 0
$$

## Optimisation Algorithm 

$$
\begin{aligned}
	&z \sim \mathcal{N}(0, 1); \epsilon_i ~ \sim \mathcal{N}(0, \sigma_{\epsilon_i}) \\
\end{aligned}
$$

## Optimisation Algorithm 

$$
\begin{aligned}
	&z \sim \mathcal{N}(0, 1); \epsilon_i ~ \sim \mathcal{N}(0, \sigma_{\epsilon_i}) \\
    \forall 0 < o \leq O: &x^o_i \sim U[-5,5] \\
\end{aligned}
$$

## Optimisation Algorithm 

$$
\begin{aligned}
	&z \sim \mathcal{N}(0, 1); \epsilon_i ~ \sim \mathcal{N}(0, \sigma_{\epsilon_i}) \\
    \forall 0 < o \leq O: &x^o_i \sim U[-5,5] \\
    \forall 0 < o \leq O: &\hat{y^o} = (\gamma_1 z + \epsilon_1) x^o_1 
						+ (\gamma_2 z  + \epsilon_2) x^o_2+ \beta_{12} x^o_1 x^o_2 \\ 
     &- \frac{1}{2}\delta_1 {x^o}^2_1 - \frac{1}{2}\delta_2 {x^o}^2_2 \\
     \forall 0 < o \leq O: &y^o \sim \mathcal{N}(\hat{y^o}, \sigma_{\nu}) \\
	 &y^{max} = max(y^1, y^2, ..., y^O)  \\
\end{aligned}
$$

## Optimisation Algorithm 

$$
\begin{aligned}
	&z \sim \mathcal{N}(0, 1); \epsilon_i ~ \sim \mathcal{N}(0, \sigma_{\epsilon_i}) \\
    \forall 0 < o \leq O: &x^o_i \sim U[-5,5] \\
    \forall 0 < o \leq O: &\hat{y^o} = (\gamma_1 z + \epsilon_1) x^o_1 
						+ (\gamma_2 z  + \epsilon_2) x^o_2+ \beta_{12} x^o_1 x^o_2 \\ 
     &+ \frac{1}{2}\delta_1 {x^o}^2_1 + \frac{1}{2}\delta_2 {x^o}^2_2 \\
     \forall 0 < o \leq O: &y^o \sim \mathcal{N}(\hat{y^o}, \sigma_{\nu}) \\
	 &y^{max} = max(y^1, y^2, ..., y^O)  \\
     &\text{if } y^o = y^{max} \text{then } y = y^o, x_i = x^o_i
\end{aligned}
$$

## Simulation parameters

$$
\begin{aligned}
\beta_1 &= \beta_2 = \beta_0 = 0 \\
\delta_1 &= \delta_2 = 1 \\
\sigma_{\epsilon_1}  &= \sigma_{\epsilon_1} = \sigma_{\nu} = 1 \\
\beta_{12} &= .25 \\
\gamma_1 &= .33, \gamma_2 = -.33
\end{aligned}
$$

## Is there sufficient variation in optimality?

```{r descriptives, eval = FALSE, include=FALSE}
source("R/sample_descriptives.R", echo = FALSE)
saveRDS(plot_opt, "R/plot_opt.RDS")
saveRDS(plot_xx, "R/plot_xx.RDS")
saveRDS(plot_xz, "R/plot_xz.RDS")
```

```{r plot_opt}
plot_opt = readRDS("R/plot_opt.RDS")
print(plot_opt)
```

## Are the simulations representative of a typical study?

```{r plot_xx}
plot_xx = readRDS("R/plot_xx.RDS")
print(plot_xx)
```

## Are the simulations representative of a typical study?

```{r plot_xz}
plot_xz = readRDS("R/plot_xz.RDS")
print(plot_xz)
```

## Type 1 Error

![Type 1 Error](figure-latex/main_null_plot.pdf)

## Power

![Power](figure-latex/main_true_plot.pdf)

# Literature
## References

\tiny
